---
title: "Class 18 code"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```




```{r}

library(SDS100)


```






$\\$




##  Normal distributions



### Normal density curves


We can plot a normal density curve using the `dnorm(x_vals, mu, sigma)` function.

The arguments to the `dnorm()` function are: 

- `x_vals`: a vector of x-values that we will get the corresponding y-values. We
can use the `seq(start_val, end_val, length.out)` argument to create these x-vals.

- `mu`: the mean of the normal density curve

- `sigma`: the standard deviation of the normal density curve



Try plotting a normal density curve with a mean of 20 and a standard deviation of 3.


```{r}

x_vals <- seq(7, 33, length.out = 1000)

y_vals <- dnorm(x_vals, 20, 3)

plot(x_vals, y_vals, type = "l")

```




$\\$




### Normal cumulative distribution functions 


We can get the probability of getting a random value less than `x` from a normal distribution using the `pnorm(x, mu, sigma)` function.

The arguments to the `pnorm()` function are: 

- `x`: the value such that P(X < x)

- `mu`: the mean of the normal density curve

- `sigma`: the standard deviation of the normal density curve


Try to get the probability of getting a value less than 15 from a normal distribution with a mean of 20 and a standard deviation of 3.


```{r}

pnorm(15, 20, 3)


# library(mosaic)
# xpnorm(15, 20, 3)


```




$\\$




### Normal quantile function 


We can get the quantile value from a normal distribution using the `qnorm()` function. To get a quantile value, we give probability value `p` that is between 0 and 1. The function returns the value `x` such that `P(X < x) = p`. 


The`qnorm(p, mu, sigma)` function has the following arguments: 

- `p`: a value between 0 and 1 such that `P(X < x) = p`

- `mu`: the mean of the normal density curve

- `sigma`: the standard deviation of the normal density curve


Try to get the quantile value such that 30% of a normal normal distribution with a mean of 20 and a standard deviation of 3 is less than the value returned. 


```{r}

qnorm(.3, 20, 3)


# xqnorm(.3, 20, 3)


``` 





$\\$




## The standard normal distribution 


The standard normal distribution is a normal distribution with a mean of 0 and a standard deviation of 1. 

We can transform any arbitrary normally distributed random variable X to a standard normal distribution using: 

$$Z = \frac{(X - \mu)}{\sigma}$$

Conversely, we can transform any standard normally distributed random variable Z into an arbitrary normally distributed random variable X using: 


$$X = \mu + \sigma \cdot Z$$

Let's explore this by generating a 10,000 normal random numbers in R using the `rnorm(10000, mu, sigma)` function and then transforming them to a standard normal distribution. Let's use a mean of 10 and a standard deviation of 3 for the numbers we generate. 


```{r}

# generate normally distributed random data with mean of 10 and a standard deviation of 3
rand_nums <- rnorm(10000, 10, 3)


# visualize the data
hist(rand_nums, breaks = 100)


# transform the data to a standard normal and plot it
transformed_random_data <- (rand_nums - 10)/3
hist(transformed_random_data, breaks = 100)


# look at the mean and standard deviation of the transformed data
mean(transformed_random_data)
sd(transformed_random_data)


```



$\\$


Let's now generate standard normal data and transform it to a normal distribution with mean of 30 and a standard deviation of 5.


```{r}

rand_nums <- rnorm(10000, 0, 1)

# visualize the data
hist(rand_nums, breaks = 100)


# transform the data to a standard normal and plot it
transformed_random_data <-  5 * rand_nums   + 30 
hist(transformed_random_data, breaks = 100)


# look at the mean and standard deviation of the transformed data
mean(transformed_random_data)
sd(transformed_random_data)


```



$\\$




## The Central Limit Theorem


The central limit theorem (CLT) establishes that, for identically distributed independent samples, the sample mean tends towards a normal distribution .

Let's explore this by generating random data in R that is right skewed (using the `r`exp()` function). We can then show that the sampling distribution of sample means is normally distributed. 


```{r}

library(SDS100)


# generate 100 points from an exponential distribution 
one_sample <- rexp(100)
hist(one_sample)


# take the mean of these points
mean(one_sample)



# create a sampling distribution with 10,000 statistics in it
sampling_dist <- do_it(10000) * {
  
  one_sample <- rexp(100)
  mean(one_sample)

}



# visualize the sampling distribution
hist(sampling_dist, breaks = 100)



```



$\\$




## Inference using normal distributions 



### Hypothesis test for a single proportion with a known SE


Do goalies guess the direction of a penalty shot less than 50% of the time? 


From 1982 to 1994 there were 128 penalty shots in the World Cup.  
Goal keepers correctly guessed the direction 41% of the time with SE* = 0.043                             



Step 1 

$H_0$: 
$H_A$:



```{r}



# Step 2: 
(z <- (.41 - .5)/.043)



# steps 3 and 4
pnorm(z, 0, 1)



# step 5




```



$\\$



### CI for a single mean with a known SE

A dataset of 200 ICU patients found that the average age of patients was 57.55 with a standard error of SE = 1.42

Use the normal distribution to compute a 90%, 95% and 99% CIs for the average age of patients in the ICU


```{r}


57.55 - qnorm(c(.90, .95, .975, .99, .995)) * 1.42     
57.55 + qnorm(c(.90, .95, .975, .99, .995)) * 1.42 


```

    




