---
title: "Class 10 code"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```




```{r}

library(SDS100)


```







## Bechdel test bootstrap CI 



$\\$




One more bootstrap confidence interval for the mean gross revenue of movies.





```{r}


# Let's create a confidence interval for the mean gross revenue of movies
library(fivethirtyeight)



# View(bechdel)
# get the data
gross <- na.omit(bechdel$domgross_2013)


# get number of data points in our sample



# get x-bar




# create a bootstrap distribution







# visualize the bootstrap distribution




# calculate SE*




# get the CI







```



Questions?

  1. How do we interpret the results?
  2  Do we believe these results?
  3. Are there any concerns about the analysis?
     a. Is the Bechdel data really a simple random sample from all movies?
     b. Might the revenue of movies change over time (even after adjusting for inflation)?
     c. In general, what is the population we are trying to do inference on?
  




$\\$






## Hypothesis testing with Doris and Buzz 



```{r}


# flipping random coins with rflip_count(num_flips, prob = .5)


# can you simulate the number of heads you will get for:
#  a. 20 coin flips
#  b. with a probability of heads on each flip being 0.75




# step 1: state null and alternative hypotheses
# H0: pi = 0.5
# HA: pi > 0.5


# step 2: compute observed statistic




# step 3: create the null distribution







# plot the null distribution






# step 4: calculate the p-value




# 5. decision?


```





$\\$









##  Testing the Bechdel test


```{r}




# step 1: state null and alternative hypotheses





# step 2: compute observed statistic





# step 3: create the null distribution









# plot the null distribution






# step 4: calculate the p-value







# 5. decision?





```




$\\$






##  Hypothesis testing with Joy  


```{r}




# step 1: state null and alternative hypotheses





# step 2: compute observed statistic




# step 3: create the null distribution








# plot the null distribution







# step 4: calculate the p-value






# 5. decision?





```







