---
title: "Class 20 code"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```




```{r}

library(SDS100)


```



$\\$





## Hypothesis test for a single proportion


Does the AstraZeneca vaccine cause blood clots? 


A study found that 79 people experienced clots after receiving a first vaccine dose.More than 20 million AstraZeneca vaccines doses had been administered across the UK by the end of March.


About four people in a million would normally be expected to develop this
particular kind of blood clot - though the fact they are so rare makes the 
usual rate hard to estimate.


### Step 1: 





```{r}


# Step 2: calculate the observed z-statistic







# Step 2b: Are these conditions met? 




# Step 3: visualize the null distribution









# Step 4: get the p-value




# Step 5: conclusion!




```




$\\$




## CI for a single proportion


What is the probability of having a blood clot if you take the AstraZeneca vaccine? 

```{r}






```



Is it likely you will get a blood clot? 





$\\$




## Confidence interval for a single proportion using the bootstrap


```{r}


library(tictoc)


has_clot <- rep(TRUE, 79)
no_clot <- rep(FALSE, (20 * 10^6) - 79)

data_vec <- c(has_clot, no_clot)





# Time how long it takes to create 10 bootstrap statistics... 
tic()








toc()






```





$\\$





## Confidence interval for a single mean

How many birds to cats kill? 


A study by Loyd et al (20213) in Biological Conservation, used KittyCams to
record all activity of n = 55 domestic cats that hunt outdoors.

The video footage showed that the mean number of kills per week for these cats
was 2.4 with a standard deviation of 1.51.

Find and interpret a 99% confidence interval for the mean number of kills per
week by US household cats that hunt outdoors.


```{r}








```







$\\$







##  Hypothesis test for a single mean


It is recommended that adults sleep at least 8 hours a night

A Statistics professor asked 12 undergraduate students how much sleep they were
getting and found the average was 6.2 hours with a standard deviation of 1.7
hours.

Assuming this is representative of all students in a Statistics class, does this
provide evidence that students in the class are not getting enough sleep on
average?


```{r}









```









