---
title: "Class 10 code"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```









## Normal quantiles

We use quantiles from a standard normal distribution to create confidence intervals for different confidence levels. 

For a given confidence level *C*, we want to find the quentile value q* such that C percent of the area of a standard normal distribution is between -q* and q*. 


```{r}

# install.packages("mosaic")


quantile_values <- mosaic::xqnorm(c(.025, .975))


quantile_values




```


$\\$



##  Simple example of sampling with replacement



```{r}


library(SDS100)


# simple vector
my_sample <- c(3, 1, 4, 1, 5, 9)


# resampling values in the vector with replacement
boot_sample  <-  sample(my_sample,  6,  replace = TRUE)


```



$\\$






##  Simple fake data bootstrap example


```{r}

# simple bootstrap example

my_sample <- c(21, 29, 25, 19, 24, 22, 25, 26, 25, 29)


# calculate the x-bar statistic
obs_mean <- mean(my_sample)


# create the bootstrap distribution
bootstrap_dist <-  do_it(10000) * {    
  
  curr_boot <- sample(my_sample , 10, replace = TRUE)  
  mean(curr_boot)  
  
}


# visualize the bootstrap distribution
hist(bootstrap_dist, 100)


# calculate the standard error
SE_boot <- sd(bootstrap_dist)


# 95% confidence interval
CI_lower <-  obs_mean  - 2 * SE_boot 
CI_upper <-  obs_mean  + 2 * SE_boot 


c(CI_lower, CI_upper)



```




$\\$




## Confidence interval example 1: One true love? 


A survey asked 2625 people whether they agreed with the statement “There is only one true love for each person”.

1812 of the respondents disagreed.

Compute a 90% confidence interval for the proportion who disagreed.





```{r}


# create the data set

belives_in_true_love <- c(rep(FALSE, 1812), rep(TRUE, 2625 - 1812))


# The actual proportion of people who believe in true love
# Note: we can use the mean function here to get a proportion
# Alternatively we could use:  get_proportion(data, categorory_name)






# We can use the sample(data_vec, n) to get a sample of length n:




# Create a bootstrap distribution in R










# visualize the bootstrap distribution





# get the bootstrap standard error SE*





# 90% confidence interval






```







$\\$






## Confidence interval example 2: Average movie revenue

Let's use the Bechdel data to calculate a confidence interval fro the mean revenue that movies make. 


```{r}

library(fivethirtyeight)

revenue <- na.omit(bechdel$domgross_2013)


# get the observed statistic




# Create a bootstrap distribution in R








# visualize the bootstrap distribution




# get the bootstrap standard error SE*




# 95% confidence interval






```





$\\$





## Are there any other questions you want to explore on your own???


For example, can you create a confidence interval for the proportion of movies that pass the Bechdel test? 





