---
title: "Class 19 code"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```




```{r}

library(SDS100)


```



$\\$





##  Hypothesis test for a single proportion with a known SE


Do goalies guess the direction of a penalty shot less than 50% of the time? 


From 1982 to 1994 there were 128 penalty shots in the World Cup.  
Goal keepers correctly guessed the direction 41% of the time With SE* = 0.043                             



### Step 1: 

$H_0$: 
$H_A$:



### Step 2-5: 

```{r}


# step 2: calculate the z-statistic




# step 3-4: use the pnorm function to get the p-value



```


Make a decision!





$\\$





##  A formula for the SE for proportions


The standard error for a single proportion is given by the formula:



$$SE = \sqrt{\frac{\pi(1-\pi)}{n}}   $$

When running a hypothesis test for a single proportion, we assume that $\pi$ is equal to the value specified by the null hypothesis ($\pi_0$) so we can calculate the standard error as:


$$SE = \sqrt{\frac{\pi_0(1-\pi_0)}{n}}$$

Let's calculate the SE for the soccer example...


```{r}




```






$\\$





## Another example of a hypothesis test for a single proportion


Does the AstraZeneca vaccine cause blood clots? 


A study found that 79 people experienced clots after receiving a first vaccine dose. 
More than 20 million AstraZeneca vaccines doses had been administered across the UK by the end of March.


About four people in a million would normally be expected to develop this
particular kind of blood clot - though the fact they are so rare makes the 
usual rate hard to estimate.


### Step 1: 





```{r}


# Step 2: calculate the observed z-statistic







# Step 2b: Are these conditions met? 




# Step 3: visualize the null distribution









# Step 4: get the p-value




# Step 5: conclusion!




```




$\\$




## CI for a single proportion


What is the probability of having a blood clot if you take the AstraZeneca vaccine? 

```{r}


# calculate the SE



# get the critical value z* for a 95% confidence interval



# create the confidence interval



```



Is it likely you will get a blood clot? 





$\\$




## Confidence interval for a single proportion using the bootstrap


```{r}


library(tictoc)


has_clot <- rep(TRUE, 79)
no_clot <- rep(FALSE, (20 * 10^6) - 79)

data_vec <- c(has_clot, no_clot)





# Time how long it takes to create 10 bootstrap statistics... 
tic()








toc()






```











