---
title: "Class 19 code"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```




```{r}

library(SDS100)


```



$\\$


## Review of normal distribution functions 

Let's review functions for normal distributions by examining a normal distribution that has a mean of 3 and a standard deviation of 20.



```{r}

# plot a normal distribution 
x_vals <- seq(7, 33, length.out = 1000)




# Calculate P(X < 15) 




# Get the quantile value q* such that there is a 20% probability of getting a value less than q*




# generate 200 random numbers




```




$\\$





## The Central Limit Theorem


The central limit theorem (CLT) establishes that, for identically distributed independent samples, the sample mean tends towards a normal distribution .

Let's explore this by generating random data in R that is right skewed (using the `r`exp()` function). We can then show that the sampling distribution of sample means is normally distributed. 


```{r}

library(SDS100)


# generate 100 points from an exponential distribution 
one_sample <- rexp(100)
hist(one_sample)


# take the mean of these points





# create a sampling distribution with 10,000 statistics in it







# visualize the sampling distribution



```






$\\$







##  Hypothesis test for a single proportion with a known SE


Do goalies guess the direction of a penalty shot less than 50% of the time? 


From 1982 to 1994 there were 128 penalty shots in the World Cup.  
Goal keepers correctly guessed the direction 41% of the time With SE* = 0.043                             



### Step 1: 

H0: 
HA:



### Step 2: 

```{r}




```








## CI for a single mean with a known SE

A data set of 200 ICU patients found that the average age of patients was 57.55 with a standard error of SE = 1.42

Use the normal distribution to compute a 90%, 95% and 99% CIs for the average age of patients in the ICU


```{r}


  


```




$\\$




## Hypothesis test for a single proportion


Does the AstraZeneca vaccine cause blood clots? 


A study found that 79 people experienced clots after receiving a first vaccine dose. 
More than 20 million AstraZeneca vaccines doses had been administered across the UK by the end of March.


About four people in a million would normally be expected to develop this
particular kind of blood clot - though the fact they are so rare makes the 
usual rate hard to estimate.


### Step 1: 





```{r}


# Step 2: calculate the observed z-statistic







# Step 2b: Are these conditions met? 




# Step 3: visualize the null distribution









# Step 4: get the p-value




# Step 5: conclusion!




```




$\\$




## CI for a single proportion


What is the probability of having a blood clot if you take the AstraZeneca vaccine? 

```{r}






```



Is it likely you will get a blood clot? 





$\\$




## Confidence interval for a single proportion using the bootstrap


```{r}


library(tictoc)


has_clot <- rep(TRUE, 79)
no_clot <- rep(FALSE, (20 * 10^6) - 79)

data_vec <- c(has_clot, no_clot)





# Time how long it takes to create 10 bootstrap statistics... 
tic()








toc()






```











