---
title: "Class 12 code"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```




```{r}

library(SDS100)


```





## Lie detector hypothesis test


$\\$


```{r}



# step 1: state null and alternative hypotheses





# step 2: compute observed statistic 
# (out of 48 questions, made correct detections 31 times)





# step 3: create the null distribution






# plot the null distribution







# step 4: calculate the p-value





# 5. decision?





```





What would have happened if we had tested whether lie detector tests get it correct more than 50% of the time? 








$\\$






## Should you choose B on a multiple choice test? 

Multiple-choice questions on Advanced Placement exams have five options: A, B, C, D, and E. A random sample of the correct choice on 400 multiple-choice questions on a variety of AP shows that B was the correct answer 90 of the 400 questions. Does this provide evidence that B occurs at a different rate than what is expected if each question is equally likely? 

Please run a hypothesis test to find out!



$\\$



```{r}



# step 1: state null and alternative hypotheses




# step 2: compute observed statistic




# step 3: create the null distribution







# plot the null distribution








# step 4: calculate the p-value





# 5. decision?





# FYI: You don't need to look at the actual data to solve this problem, 
# but if you'd like it can be loaded using:
#  APMultipleChoice <- Lock5Data::APMultipleChoice



# Bonus, can you create a 95% CI for the proportion of answers that are B? 




```






$\\$




