---
title: "Class 12 code"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```




```{r}

library(SDS100)


```





## Lie detector hypothesis test


$\\$


```{r}



# step 1: state null and alternative hypotheses





# step 2: compute observed statistic 
# (out of 48 questions, made correct detections 31 times)





# step 3: create the null distribution






# plot the null distribution







# step 4: calculate the p-value





# 5. decision?





```





What would have happened if we had tested whether lie detector tests get it correct more than 50% of the time? 








$\\$






## Should you choose B on a multiple choice test? 

Multiple-choice questions on Advanced Placement exams have five options: A, B, C, D, and E. A random sample of the correct choice on 400 multiple-choice questions on a variety of AP shows that B was the correct answer 90 of the 400 questions. Does this provide evidence that B occurs at a different rate than what is expected if each question is equally likely? 

Please run a hypothesis test to find out!



$\\$



```{r}



# step 1: state null and alternative hypotheses




# step 2: compute observed statistic




# step 3: create the null distribution







# plot the null distribution








# step 4: calculate the p-value





# 5. decision?





# FYI: You don't need to look at the actual data to solve this problem, 
# but if you'd like it can be loaded using:
#  APMultipleChoice <- Lock5Data::APMultipleChoice



# Bonus, can you create a 95% CI for the proportion of answers that are B? 




```






$\\$







## Hypothesis testing with Doris and Buzz 

```{r}


# flipping random coins with rflip_count(num_flips, prob = .5)


# can you simulate the number of heads you will get for:
#  a. 20 coin flips
#  b. with a probability of heads on each flip being 0.75

rflip_count(20, .75)


# step 1: state null and alternative hypotheses
# H0: pi = 0.5
# HA: pi > 0.5


# step 2: compute observed statistic
(buzz_phat <- 15/16)



# step 3: create the null distribution

null_dist <- do_it(10000) * {
  
  rflip_count(16, .5)/16
  
}


hist(null_dist, breaks = 100)
abline(v = buzz_phat, col = "red")



# step 4: calculate the p-value

pnull(buzz_phat, null_dist, lower.tail = FALSE)



# 5. decision?






```





$\\$






##  Hypothesis testing with Joy  


```{r}




# step 1: state null and alternative hypotheses




# step 2: compute observed statistic



# step 3: create the null distribution









# step 4: calculate the p-value





# 5. decision?






```








